Open Data, Open Source and Open Standards in chemistry: \\ The Blue Obelisk five years on

### 2011-08-17 paper.tex
Final edits responding to reviewer's comments

Despite this positive news, little chemical data has become
available from the traditional chemical fields of organic,
inorganic, solid state chemistry. Table~2 lists a few notable
exceptions: Metamolecular's Chempedia~\cite{Chempedia} (a discontinued substance registration
and naming service whose raw data can still be accessed),
CrystalEye~\cite{WebCrystalEye},
and the Open Notebook Science Solubility
data~\cite{ONS2010}. There is also data available using licenses
not compatible with the Panton Principles, but where the user
is allowed to modify and redistribute the data. A new data
set in this category is the data from the ChEMBL database,
which is available under the Creative Commons Share-Alike
Attribution license~\cite{Overington2009}. The RSC ChemSpider database~\cite{chemspider}, although
not fully Open, also hosts Open Data; for example, spectral data when deposited
can be marked as Open.

@@

The Blue Obelisk has created a repository of key chemical
data in a machine-readable format~\cite{BODR}.
The BODR focuses on data that is commonly required for
chemistry software, and where there is a need to ensure that values
are standard between codes. Examples are atomic
masses and conversions between physical constants. These data
can be used by others for any purpose (for example, for entry into
Wikipedia or use in in-house software), and should lead
to an enhancement in the quality of community reference data.
The Blue Obelisk provides also a complementary project, the Chemical
Structure Repository~\cite{BODR}. It aims to provide 3D coordinates, InChi code and several
physico-chemical descriptors for a set of 570 organic compounds.

@@

  \subsection*{Table 2 - Open Data in chemistry.}
    Overview of major open chemical data available under a license or waiver
    compatible with the Panton Principles.
    \par \mbox{}
    \par
    \mbox{
\begin{tabular}{|l|c|l|}
  %% \hline \multicolumn{3}{|c|}{My Table}\\ \hline
  \hline \textbf{Name} & \textbf{License/Waiver} & \textbf{Description} \\ \hline
  \textbf{Chempedia} & CC0 & Crowd-sourced chemical names (project discontinued) \\ \hline
  \textbf{CrystalEye}     & PPDL & Crystal structures from primary literature \\ \hline
  \textbf{ONS Solubility} & CC0 & Solubility data for various solvents \\ \hline
  \textbf{Reaction Attempts} & CC0 & Data on successful and unsuccessful reactions \\ \hline
\end{tabular}
      }

>>>>

Despite this positive news, little chemical data compatible
with these principles has become
available from the traditional chemistry fields of organic,
inorganic, and solid state chemistry. Table~2 lists a few notable
exceptions, some of which are discussed further below.
There is also data available using licenses
not compatible with the Panton Principles, but where the user
is allowed to modify and redistribute the data. A new data
set in this category is the data from the ChEMBL database~\cite{Overington2009},
which is available under the Creative Commons Share-Alike
Attribution license. The RSC ChemSpider database~\cite{chemspider}, although
not fully Open, also hosts Open Data; for example, spectral data when deposited
can be marked as Open.

@@

The Blue Obelisk has created a repository of key chemical
data in a machine-readable format~\cite{BODR}.
The BODR focuses on data that is commonly required for
chemistry software, and where there is a need to ensure that values
are standard between codes. Examples are atomic
masses and conversions between physical constants. These data
can be used by others for any purpose (for example, for entry into
Wikipedia or use in in-house software), and should lead
to an enhancement in the quality of community reference data.
The Blue Obelisk provides also a complementary project, the Chemical
Structure Repository~\cite{BODR}. It aims to provide 3D coordinates, InChIs and several
physico-chemical descriptors for a set of 570 organic compounds.

@@

  \subsection*{Table 2 - Open Data in chemistry.}
    Overview of major open chemical data available under a license or waiver
    compatible with the Panton Principles.
    \par \mbox{}
    \par
    \mbox{
\begin{tabular}{|l|c|l|}
  %% \hline \multicolumn{3}{|c|}{My Table}\\ \hline
  \hline \textbf{Name} & \textbf{License/Waiver} & \textbf{Description} \\ \hline
  \textbf{Chempedia}~\cite{Chempedia} & CC0 & Crowd-sourced chemical
names (project discontinued but data still available) \\ \hline
  \textbf{CrystalEye}     & PPDL & Crystal structures from primary literature \\ \hline
  \textbf{ONS Solubility} & CC0 & Solubility data for various solvents \\ \hline
  \textbf{Reaction Attempts} & CC0 & Data on successful and unsuccessful reactions \\ \hline
\end{tabular}
      }
